
http://ccb.jhu.edu/software/stringtie/dl/prepDE.py3
http://ccb.jhu.edu/software/stringtie/index.shtml?t=manual#deseq

python prepDE.py3 -i OpossumPython.txt -g opossum_gene_count_matrix.csv -t opossum_transcript_count_matrix.csv-l 200

python prepDE.py3 -i MousePython.txt -g mouse_gene_count_matrix.csv -t mouse_transcript_count_matrix.csv-l 36

#-i is the input file that lists the paths to the GTF files from stringtie
# -g where to output the gene count matrix [default: gene_count_matrix.csv]
# -t where to output the transcript count matrix [default: transcript_count_matrix.csv]
#-l the average read length [default: 75]

#loading necessary packages
```{r}
library(tidyverse)
library(matrixStats)
library(biomaRt)
library(DESeq2)
library(pheatmap)
library(RColorBrewer)
```


#1 Read in and format tables
reading in output tables from HISAT-STRINGTIE using -e option followed by python script to format data
 
##Read in Gene Count table
```{r}
setwd("~/Desktop/Analyses/Stringtie_Output/ForDESeq")
countData <- as.matrix(read.csv("opossum_gene_count_matrix_comb.csv", row.names="gene_id"))
```

#Make a data table to describe data for opossum
```{r}
samples <- colnames(countData) #makes a vector of all the samples
sampleConditions_md <- c("2M", "2M", "2M", "3M", "3M", "3M", "3M", "4M", "4M", "4M", "AD", "AD", "AD", "AD", "AD", "AD", "2M", "AD", "1M", "1M")
#makes a vector of the ages for each sample
colData <- data.frame(Samples = samples, condition = sampleConditions_md, stringsAsFactors = FALSE)
#combines vectors into a dataframe
rownames(colData) <- colData$Samples #makes the sample names into the row names
```

```{r}
#Check all sample IDs in colData are also in CountData and match their orders
all(rownames(colData) %in% colnames(countData))
countData <- countData[, rownames(colData)]
all(rownames(colData) == colnames(countData))
```

##Create a DESeqDataSet from count matrix and labels
```{r}
#make DESeq object
dds <- DESeqDataSetFromMatrix(countData = countData, colData = colData, design = ~ condition)
dds <- DESeq(dds) #runs DESeq analysis

res <- results(dds, pAdjustMethod = 'fdr')
```

```{r}
#all from my notes from genomics
res_data <- as.data.frame(res)#makes it into a proper data frame
res_data_rows <- rownames_to_column(res_data, var = "gene") %>% arrange(padj)
#makes the gene names into a column instead of the row names


resOrdered <- res_data[order(res_data$padj),] #reorders by p value 

sum(res_data$padj < 0.05, na.rm = TRUE) #this tells you the number of occurrences with a p value below 0.05

```

```{r}
setwd("~/Desktop/Analyses/tx2gene_tables")
write.table(res_data, file = "Opossum_DEseq.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

#Extracting pairwise data
##1-2 Months

```{r}
res_1_2 <- results(dds, contrast = c("condition","1M", "2M"))
res_data_1_2 <- as.data.frame(res_1_2)#makes it into a proper data frame
sum(res_data_1_2$padj < 0.05, na.rm = TRUE) 
sum(res_data_1_2$pvalue < 0.05, na.rm = TRUE) 
```

##2-3 Months

```{r}
res_2_3 <- results(dds, contrast = c("condition","2M", "3M"))
res_data_2_3 <- as.data.frame(res_2_3)#makes it into a proper data frame
sum(res_data_2_3$padj < 0.05, na.rm = TRUE) 
sum(res_data_2_3$pvalue < 0.05, na.rm = TRUE) 
```

##3-4 Months
```{r}
res_3_4 <- results(dds, contrast = c("condition","3M", "4M"))
res_data_3_4 <- as.data.frame(res_3_4)#makes it into a proper data frame
sum(res_data_3_4$padj < 0.05, na.rm = TRUE) 
sum(res_data_3_4$pvalue < 0.05, na.rm = TRUE) 
```

## 4M- AD Months
```{r}
res_4_ad <- results(dds, contrast = c("condition","4M", "AD"))
res_data_4_ad <- as.data.frame(res_4_ad)#makes it into a proper data frame
sum(res_data_4_ad$padj < 0.05, na.rm = TRUE) 
sum(res_data_4_ad$pvalue < 0.05, na.rm = TRUE) 
```

##write tables
```{r}

write.table(res_data_1_2, file = "Opossum_DEseq_1_2M.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(res_data_2_3, file = "Opossum_DEseq_2_3M.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(res_data_3_4, file = "Opossum_DEseq_3_4M.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(res_data_4_ad, file = "Opossum_DEseq_4M_AD.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```



```{r}
plotMA(dds, ylim=c(-2,2), alpha = 0.05, colNonSig = "gray60", colSig = "blue")

vst <- vst(dds, blind = TRUE)

Md_vst_PCA_data <- plotPCA(vst, returnData = TRUE)
Md_vst_PCA <- plotPCA(vst)
Md_vst_PCA

ggplot(Md_vst_PCA_data, aes(x = PC1, y = PC2, color = factor(condition))) + 
geom_point(size =3, aes(fill=factor(condition))) + geom_point(size =3) + scale_color_manual(values=c("#FFCCCC", "#FF9999", "#FF6699", "#CC3399", "#990066")) + ggtitle("PCA Opossum Sub") + theme_classic()
```




#Hierarchical clustering

```{r}
sampleDists <- dist(t(assay(vst))) #gets "distance" between the samples following vst transformation
sampleDistMatrix <- as.matrix(sampleDists) #turns that into a usable matrix
rownames(sampleDistMatrix) <- paste(vst$condition) #adds the condition (age) as a row name to the matrix
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
pheatmap(sampleDistMatrix,
         clustering_distance_rows=sampleDists,
         clustering_distance_cols=sampleDists,
         col=colors)
#makes a heat map of distances along with the hierarchical clustering diagram
```

```{r}
#to make a hierarchical clustering dendrogram without the heatmap
#still using vst transformed data

rv <- rowVars(assay(vst))
o <- order(rv,decreasing=TRUE)
dists <- dist(t(assay(vst)[head(o,500),]))

hc <- hclust(dists)
plot(hc, labels=vst$sample)
```

#Making normalized count and means tables
```{r}
countData_df <- as.data.frame(countData)
countData_df$Gene_ID <- rownames(countData_df)
countData_df <- separate(data = countData_df, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")
```

```{r}
countData_df$Avg2M <- rowMeans(countData_df[ , c(1:3, 17)])
countData_df$Avg1M <- rowMeans(countData_df[ , c(19:20)])
countData_df$Avg3M <- rowMeans(countData_df[ , c(4:7)])
countData_df$Avg4M <- rowMeans(countData_df[ , c(8:10)])
countData_df$AvgAD <- rowMeans(countData_df[ , c(11:16, 18)])
```

```{r}
write.table(countData_df, file = "DESeq_Counts.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

```{r}
counts_norm <- counts(dds, normalized=T)
counts_norm <- as.data.frame(counts_norm)
counts_norm$Gene_ID <- rownames(counts_norm)
counts_norm <- separate(data = counts_norm, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")
```

```{r}
counts_norm$Avg1M_Norm <- rowMeans(counts_norm[ , c(19:20)])
counts_norm$Avg2M_Norm <- rowMeans(counts_norm[ , c(1:3, 17)])
counts_norm$Avg3M_Norm <- rowMeans(counts_norm[ , c(4:7)])
counts_norm$Avg4M_Norm <- rowMeans(counts_norm[ , c(8:10)])
counts_norm$AvgAD_Norm <- rowMeans(counts_norm[ , c(11:16, 18)])

#counts_norm$diff1M2M <- abs(counts_norm$Avg2M - counts_norm$Avg1M)
#counts_norm$div1M2M <- abs(counts_norm$Avg2M/counts_norm$Avg1M)
```

```{r}
Means_table <- subset(counts_norm[ ,c(21:27)])
```

```{r}
Means_table$Avg1_2_Norm <- rowMeans(Means_table[ , c(3:4)])

Means_table$Avg2_3_Norm <- rowMeans(Means_table[ , c(4:5)])

Means_table$Avg3_4_Norm <- rowMeans(Means_table[ , c(5:6)])

Means_table$Avg4_AD_Norm <- rowMeans(Means_table[ , c(6:7)])
```

```{r}
write.table(Means_table, file = "DESeq_NormMeansTable.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```


## Subsetting based on fold change as well 

#Using 4-fold as cutoff

```{r}
res_data_1_2_sig <- filter(res_data_1_2, padj <= 0.05 & abs(log2FoldChange) >= 2)
res_data_2_3_sig <- filter(res_data_2_3, padj <= 0.05 & abs(log2FoldChange) >= 2)
res_data_3_4_sig <- filter(res_data_3_4, padj <= 0.05 & abs(log2FoldChange) >= 2)
res_data_4_Ad_sig <- filter(res_data_4_ad, padj <= 0.05 & abs(log2FoldChange) >= 2)

sum(res_data_1_2$padj <= 0.05 & abs(res_data_1_2$log2FoldChange) >= 2, na.rm = TRUE)
sum(res_data_2_3$padj <= 0.05 & abs(res_data_2_3$log2FoldChange) >= 2, na.rm = TRUE)
sum(res_data_3_4$padj <= 0.05 & abs(res_data_3_4$log2FoldChange) >= 2, na.rm = TRUE)
sum(res_data_4_ad$padj <= 0.05 & abs(res_data_4_ad$log2FoldChange) >= 2, na.rm = TRUE)
```


```{r}
#to add gene names as a column and separate the ID from the common name
res_data_1_2_sig$Gene_ID <- rownames(res_data_1_2_sig)
res_data_1_2_sig <- separate(data = res_data_1_2_sig, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

#to add gene names as a column and separate the ID from the common name
res_data_2_3_sig$Gene_ID <- rownames(res_data_2_3_sig)
res_data_2_3_sig <- separate(data = res_data_2_3_sig, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

res_data_3_4_sig$Gene_ID <- rownames(res_data_3_4_sig)
res_data_3_4_sig <- separate(data = res_data_3_4_sig, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

res_data_4_Ad_sig$Gene_ID <- rownames(res_data_4_Ad_sig)
res_data_4_Ad_sig <- separate(data = res_data_4_Ad_sig, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")
```


```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("condition")])
rownames(df) <- colnames(assay(rld)[select,])
pheatmap(assay(vst)[select,], cluster_rows=FALSE,
         cluster_cols=FALSE, annotation_col=df)
pheatmap(assay(rld)[select,], cluster_rows=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```

#----
#make combined table with sig genes and normalized means
```{r}
count_data_1_2_sig <- left_join(res_data_1_2_sig, Means_table, by = "Gene_ID")

count_data_2_3_sig <- left_join(res_data_2_3_sig, Means_table, by = "Gene_ID")

count_data_3_4_sig <- left_join(res_data_3_4_sig, Means_table, by = "Gene_ID")

count_data_4_Ad_sig <- left_join(res_data_4_Ad_sig, Means_table, by = "Gene_ID")
```


```{r}
ggplot(data = count_data_1_2_sig, aes(x = Avg1M_Norm, y = Avg2M_Norm)) + geom_point()

ggplot(data = count_data_1_2_sig, aes(x = Avg1M_Norm, y = Avg2M_Norm)) + 
  geom_point() +
  xlim(0, 300000) +
  ylim(0, 300000)

ggplot(data = count_data_1_2_sig, aes(x = Avg1M_Norm, y = Avg2M_Norm)) + 
  geom_point() +
  xlim(0, 5000) +
  ylim(0, 5000)

ggplot(data = count_data_1_2_sig, aes(x = Avg1M_Norm, y = Avg2M_Norm)) + 
  geom_point() +
  xlim(0, 50) +
  ylim(0, 50)

```

#Filter based on expression means
```{r}
count_data_1_2_sig_50 <- filter(count_data_1_2_sig, Avg1_2_Norm >= 50)
count_data_2_3_sig_50 <- filter(count_data_2_3_sig, Avg2_3_Norm >= 50)
count_data_3_4_sig_50 <- filter(count_data_3_4_sig, Avg3_4_Norm >= 50)
count_data_4_Ad_sig_50 <- filter(count_data_4_Ad_sig, Avg4_AD_Norm >= 50)
```


```{r}
ggplot(data = count_data_1_2_sig, aes(x = Avg1M_Norm, y = Avg2M_Norm)) + 
  geom_point() +
  xlim(0, 100) +
  ylim(0, 100) + 
  geom_hline(yintercept=20, linetype = "dashed") + 
  geom_vline(xintercept=20, linetype = "dashed") + 
  theme_classic()
```

#Making count tables for graphing
```{r}
res_data_1_2$Gene_ID <- rownames(res_data_1_2)
res_data_1_2 <- separate(data = res_data_1_2, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

res_data_2_3$Gene_ID <- rownames(res_data_2_3)
res_data_2_3 <- separate(data = res_data_2_3, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

res_data_3_4$Gene_ID <- rownames(res_data_3_4)
res_data_3_4 <- separate(data = res_data_3_4, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

res_data_4_ad$Gene_ID <- rownames(res_data_4_ad)
res_data_4_ad <- separate(data = res_data_4_ad, col = Gene_ID, into = c("Gene_ID", "Gene_Name"), sep = "\\|")

count_data_1_2 <- left_join(res_data_1_2, Means_table, by = "Gene_ID")
count_data_2_3 <- left_join(res_data_2_3, Means_table, by = "Gene_ID")
count_data_3_4 <- left_join(res_data_3_4, Means_table, by = "Gene_ID")
count_data_4_Ad <- left_join(res_data_4_ad, Means_table, by = "Gene_ID")

```

```{r}
ggplot(data = count_data_1_2, aes(x = log2(Avg1M_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  theme_classic()

ggplot(data = count_data_2_3, aes(x = log2(Avg2M_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  theme_classic()

ggplot(data = count_data_3_4, aes(x = log2(Avg3M_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  theme_classic()

ggplot(data = count_data_4_Ad, aes(x = log2(Avg4M_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  theme_classic()
```


```{r}
ggplot(data = count_data_1_2, aes(x = log2(Avg1_2_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  geom_vline(xintercept = 5.644, color = "blue") +
  theme_classic()

ggplot(data = count_data_2_3, aes(x = log2(Avg2_3_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  geom_vline(xintercept = 5.644, color = "blue") +
  theme_classic()

ggplot(data = count_data_3_4, aes(x = log2(Avg3_4_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  geom_vline(xintercept = 5.644, color = "blue") +
  theme_classic()

ggplot(data = count_data_4_Ad, aes(x = log2(Avg4_AD_Norm), y = log2FoldChange)) + 
  geom_point() +
  geom_hline(yintercept = 2, color = "red" ) +
  geom_hline(yintercept = -2, color = "red") +
  geom_vline(xintercept = 5.644, color = "blue") +
  theme_classic()
```

```{r}
write.table(count_data_1_2_sig_50, file = "DESeq_1v2_FC4_SigMean50.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(count_data_2_3_sig_50, file = "DESeq_2v3_FC4_SigMean50.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(count_data_3_4_sig_50, file = "DESeq_3v4_FC4_SigMean50.txt", sep = "\t", row.names = FALSE, quote = FALSE)
write.table(count_data_4_Ad_sig_50, file = "DESeq_4vAd_FC4_SigMean50.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```
