#general notes:
#upMm =  genes with higher expression in mouse
#upMd = genes with higher expression in opossum
#highcons = genes highly conserved expression
#o2o = All genes used (genes with one to one orthology)




#Prep for PhastCons and PhyloP
# make bed files for gene info in R.  
```{r}
library("biomaRt")
library(dplyr)
library(data.table)
```

```{r}
upMm <- read.table("upMm_genes.txt", header = T)
upMd <- read.table("upMd_genes.txt", header = T)
highcons <- read.table("HighCons_genes.txt", header = T)
o2o <- read.table("OneToOne_genes.txt", header = T)
```

```{r}
mart=useMart("ensembl",host="https://may2021.archive.ensembl.org", dataset = "mmusculus_gene_ensembl")


tss <- getBM(attributes = c("ensembl_gene_id", "chromosome_name",
                            "transcript_start", "transcript_end", "strand",
                            "transcription_start_site", "transcript_length",
                            "ensembl_transcript_id", "external_gene_name"),
             mart = mart)
```

#filter for longest transcript (written with Delaney Farris)
```{r}
#collect unique gene names from the ensembl gene name column 
unique_genes <- unique(tss$ensembl_gene_id)
```

```{r}
#create a function longest_transcript:
#input: a subset of biomart table of a single ensembl gene, with all transcripts associated with that gene id
#output: a table with a single row associated with the longest transcript 

longest_transcript <- function(biomart_table) {
  biomart_table[order(biomart_table$transcript_length,decreasing=TRUE),][1,]
}

#generate a new empty table to add rows to
longest_transcript_table <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(longest_transcript_table) <- c('ensembl_gene_id', 'chromosome', 'start', 'end', 'strand', 'transcriptional_startsite', 'transcript_length', 'ensembl_transcript_id', 'gene_name')

#iterate through unique genes, perform "longest_transcript" function, append each gene to the table "longest_transcript_table"
for (gene in unique_genes) {
  subset_table <- tss[tss$ensembl_gene_id == gene, ]
  longest_transcript_table <- rbind(longest_transcript_table, longest_transcript(subset_table))
}

#sanity check: the final table should have as many rows as unique genes
length(unique_genes) == length(longest_transcript_table$transcript_start)
```

#make tables of biomart info for each of the gene groups
```{r}
tss_upMd <- inner_join(longest_transcript_table, upMd, by = c("ensembl_gene_id" = "Mouse_Gene"))
tss_upMm <- inner_join(longest_transcript_table, upMm, by = c("ensembl_gene_id" = "Mouse_Gene"))
tss_o2o <- inner_join(longest_transcript_table, new_o2o, by = c("ensembl_gene_id" = "Mouse_Gene"))
tss_highcons <- inner_join(longest_transcript_table, Pst_cons, by = c("ensembl_gene_id" = "Mouse_Gene"))
```


#for +-1 KB from TSS
```{r}
tss_upMd_mod <- tss_upMd[, c(2, 6, 1)]

tss_upMd_mod$up1kb <- (tss_upMd_mod$transcription_start_site - 1000) 
tss_upMd_mod$down1kb <- (tss_upMd_mod$transcription_start_site + 1000) 
```

```{r}
tss_upMd_mod <- tss_upMd_mod[, c(1, 4:5, 2:3)]
```

```{r}
tss_upMm_mod <- tss_upMm[, c(2, 6, 1)]

tss_upMm_mod$up1kb <- (tss_upMm_mod$transcription_start_site - 1000) 
tss_upMm_mod$down1kb <- (tss_upMm_mod$transcription_start_site + 1000) 
```

```{r}
tss_upMm_mod <- tss_upMm_mod[, c(1, 4:5, 2:3)]
```
 
```{r}
tss_o2o_mod <- tss_o2o[, c(2, 6, 1)]

tss_o2o_mod$up1kb <- (tss_o2o_mod$transcription_start_site - 1000) 
tss_o2o_mod$down1kb <- (tss_o2o_mod$transcription_start_site + 1000) 
```

```{r}
tss_o2o_mod <- tss_o2o_mod[, c(1, 4:5, 2:3)]
```

```{r}
tss_highcons_mod <- tss_highcons[, c(2, 6, 1)]

tss_highcons_mod$up1kb <- (tss_highcons_mod$transcription_start_site - 1000) 
tss_highcons_mod$down1kb <- (tss_highcons_mod$transcription_start_site + 1000) 
```

```{r}
tss_highcons_mod <- tss_highcons_mod[, c(1, 4:5, 2:3)]
```

```{r}
colnames(tss_upMd_mod) <- c('chrom', 'chromStart', 'chromEnd', 'TSS', 'Gene_ID')
colnames(tss_upMm_mod) <- c('chrom', 'chromStart', 'chromEnd', 'TSS', 'Gene_ID')
colnames(tss_o2o_mod) <- c('chrom', 'chromStart', 'chromEnd', 'TSS', 'Gene_ID')
colnames(tss_highcons_mod) <- c('chrom', 'chromStart', 'chromEnd', 'TSS', 'Gene_ID')
```

```{r}
tss_upMd_mod <- tss_upMd_mod[ , -c(5)]
tss_upMm_mod <- tss_upMm_mod[ , -c(5)]
tss_o2o_mod <- tss_o2o_mod[ , -c(5)]
tss_highcons_mod <- tss_highcons_mod[ , -c(5)]
```

```{r}
setwd("~/Folder")

write.table(tss_highcons_mod, "tss_highcons_mod.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

write.table(tss_o2o_mod, "tss_o2o.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

write.table(tss_upMd_mod, "tss_upMd_mod.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)

write.table(tss_upMm_mod, "tss_upMm_mod.bed",row.names = F,col.names = F, sep="\t", quote=FALSE)
```

#Phastcons analysis
#in Linux

#get phastcons scores
for phastcons
ftp> cd goldenPath/mm39/phastCons35way
ftp> get mm39.35way.phastCons.bw

```{r}
bigWigAverageOverBed mm39.phastCons35way.bw tss_upMd_mod_chrom.bed -bedOut=tss_upMd_mod_out.bed tss_upMd_mod_out.tab

bigWigAverageOverBed mm39.phastCons35way.bw tss_upMm_mod_chrom.bed -bedOut=tss_upMm_mod_out.bed tss_upMm_mod_out.tab

bigWigAverageOverBed mm39.phastCons35way.bw tss_o2o_mod_chrom.bed -bedOut=tss_o2o_mod_out.bed tss_o2o_out.tab

bigWigAverageOverBed mm39.phastCons35way.bw tss_highcons_mod_chrom.bed -bedOut=tss_highcons_mod_out.bed tss_highcons_mod_out.tab

```

#PhyloP analysis
#in Linux

#for phyloP scores
ftp> cd goldenPath/mm39/phyloP35way
ftp> get mm39.phyloP35way.bw

#phyloP
```{r}
bigWigAverageOverBed mm39.phyloP35way.bw tss_upMd_mod_chrom.bed -bedOut=tss_upMd_mod_out_phyloP.bed tss_upMd_mod_out_phyloP.tab

bigWigAverageOverBed mm39.phyloP35way.bw tss_upMm_mod_chrom.bed -bedOut=tss_upMm_mod_out_phyloP.bed tss_upMm_mod_out_phyloP.tab

bigWigAverageOverBed mm39.phyloP35way.bw tss_o2o_mod_chrom.bed -bedOut=tss_o2o_mod_out_phyloP.bed tss_o2o_out_phyloP.tab

bigWigAverageOverBed mm39.phyloP35way.bw tss_highcons_mod_chrom.bed -bedOut=tss_highcons_mod_out_phyloP.bed tss_highcons_mod_out_phyloP.tab
```

#Back in R for visualizing and stats
#PhastCon
```{r}
cons.upMm <- read.table("~/Downloads/tss_upMm_mod_out.bed", header=F)
cons.upMd <- read.table("~/Downloads/tss_upMd_mod_out.bed", header=F)
cons.o2o <- read.table("~/Downloads/tss_o2o_mod_out.bed", header=F)
cons.highcons <- read.table("~/Downloads/tss_highcons_mod_out.bed", header=F)

colnames(cons.upMm) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.upMd) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.o2o) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.highcons) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
```

```{r}
cons.o2o <- cons.o2o %>% mutate(Group ="A")
cons.upMm <- cons.upMm %>% mutate(Group ="B")
cons.upMd <- cons.upMd %>% mutate(Group ="C")
cons.highcons <- cons.highcons %>% mutate(Group ="D")
```

```{r}
cons_scores <- rbind(cons.upMm, cons.upMd)
cons_scores <- rbind(cons_scores, cons.highcons)

colnames(cons_scores) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score", "Group")
```

#Graph
```{r}
p <- ggplot(aes(x=Score, color=Group), data=cons_scores_)
p + geom_density(alpha = 0.2, size = 3) + scale_color_manual(values=c("mediumblue", "lightsteelblue", "black")) + theme_classic() + ggtitle("PhastCons_Promoter")
```

#Stats
```{r}
kruskal.test(Score ~ Group, data = cons_scores)

dunn.test_phastcons <- as.data.frame(dunn.test(cons_scores$Score, cons_scores$Group,
         method = "bonferroni", wrap = TRUE))

dunn.test_phastcons
```

#PhyloP
```{r}
cons.upMm_phyloP <- read.table("~/Downloads/tss_upMm_mod_out_phyloP.bed", header=F)
cons.upMd_phyloP <- read.table("~/Downloads/tss_upMd_mod_out_phyloP.bed", header=F)
cons.o2o_phyloP <- read.table("~/Downloads/tss_o2o_mod_out_phyloP.bed", header=F)
cons.highcons_phyloP <- read.table("~/Downloads/tss_highcons_mod_out_phyloP.bed", header=F)

colnames(cons.upMm_phyloP) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.upMd_phyloP) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.o2o_phyloP) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
colnames(cons.highcons_phyloP) <- c("Chrom", "Upstream", "Downstream", "TSS", "Score")
```


```{r}

cons.o2o_phyloP <- cons.o2o_phyloP %>% mutate(Group ="A")
cons.upMm_phyloP <- cons.upMm_phyloP %>% mutate(Group ="B")
cons.upMd_phyloP <- cons.upMd_phyloP %>% mutate(Group ="C")
cons.highcons_phyloP <- cons.highcons_phyloP %>% mutate(Group ="D")
```

```{r}
cons_scores_phyloP <- rbind(cons.upMm_phyloP, cons.upMd_phyloP)
cons_scores_phyloP <- rbind(cons_scores_phyloP, cons.highcons_phyloP)
```

#graph
```{r}
p <- ggplot(aes(x=Score, color=Group), data=cons_scores_phyloP)
p + geom_density(alpha = 0.2, lwd = 3) + scale_color_manual(values=c("mediumblue", "lightsteelblue", "black")) + theme_classic() + ggtitle("PhyloP_Promoter")
```

#stats
```{r}
kruskal.test(Score ~ Group, data = cons_scores_phyloP)

dunn.test_phylop <- as.data.frame(dunn.test(cons_scores_phyloP$Score, cons_scores_phyloP$Group,
         method = "bonferroni", wrap = TRUE))

dunn.test_phylop
```
